### 2021-04-19 MSE_paper.tex
Initial commit

\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{tikz}
\usepackage{float}
\usepackage{slashbox}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsfonts}
\usepackage{enumerate}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{bm}
\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}

@@

\begin{document}
\section{Abstract}

@@

@@

\subsection{Pure error}
The error variance estimation can be obtained in two ways. The first one is the mean square error: $\hat{\sigma}^2_{mse}=\mbox{Residual SS}/(n-p)$, i.e. the residual sum of squares divided by the corresponding number of degrees of freedom (obtained from the ANOVA decomposition, the details can be found in \cite{Draper1998}). This estimator obviously depends on the model and on the number of its parameters. The other one is model-independent `pure' error, derived from the further decomposition of the residual sum of squares into the `pure' error and `lack-of-fit' components: $\hat{\sigma}^2_{PE}=\mbox{Pure error SS}/(n-t)$, where $t$ is the number of unique treatments applied and $d=n-t$ is the pure error degrees of freedom, i.e. the number of replicated points. In other words, the error is estimated by fitting the full treatment model:

@@

@@

\section{Component criteria construction}
\label{ch::compound_criteria}
\input{compound_criterion}

@@

\section{Criteria for blocked experiments}
\label{ch::compound_blocked}
\input{compound_blocked}

\subsection{Example}
[A usual example or the case-study?]
\section{Future work}
\subsection{Multistratum experiments}
In a large number of industrial, engineering and laboratory-based experiments, either the nature of the process under study or certain technical restrictions result in the necessity of considering a multi-level unit structure. For example, a chemical process consisting of applying treatments to the material batches of different sizes at each stage;  or one or several experimental factors' values can be changed only once per a certain amount of runs whereas values of other factors are varied between runs. Therefore, different factors are applied at different levels, and randomisation is performed at each level, thus the whole process of allocating treatments to experimental units should be amended accordingly \citep{MeadGilmour2012}.

Experiments comprising a hierarchical structure of experimental units and treatments are referred to as multistratum experiments, and each stratum is defined as a level in the unit structure. Units are grouped into whole-plots, each of them divided into sub-plots, which contain a certain number of sub-sub-plots, and so on up to the smallest units --- runs of the experiment. The corresponding factors are often called ``very-hard-to-change'', ``hard-to-change'', ``easy-to-change'', ``very-easy-to-change'' factors; these labels are quite arbitrary and certainly determined by a particular experimental framework. In the case of two strata, we deal with what is called a ``split-plot'' experiment, in case of three strata a ``split-split-plot'' experiment. The names come from the terminology of agricultural experiments where such setups were originally implemented. \cite{Nelder1965analysis} introduced the notion of simple orthogonal block structures, comprising `chains' of nested and/or crossed experimental units together with the corresponding factors. Sometimes the ``split'' notation is used exceptionally for orthogonal block design structures, however, we employ it for general nested unit structures where the orthogonality property does not necessarily hold.
%\cite{Speed1982class}

Some visualisation tools allowing a comprehensive presentation of the outline of such experiments can be found in \cite{Goos2012Hasse}: Hasse diagrams, that were described extensively and in great detail by \cite{Bailey2008design}. Hasse diagrams are graphs containing the information about the experimental units, factors applied at every stratum, together with a scheme of the available degrees of freedom allocation, so that they appear to be a much useful tool providing a full and clear picture of a complicated structured experiment.

A generalised example of a multistratum experiment with a simple nested unit structure is presented in Figure \ref{Fig::Hasse}. Each node of the graph corresponds to a blocking factor and edges are drawn from one factor to another that is nested within it. The ``universe'' node U stands for the whole experiment.

From here onwards we will be considering experiments with $s$ strata in total, stratum $i$ being nested within units of the stratum $i-1$, and stratum $0$ will be seen as the whole experiment (following the notation set by \citet{Trinca2015improved}). The number of units in stratum $i$ within every unit in stratum $i-1$ is denoted by $n_i$, such that $m_{j}=\prod_{i=1}^{j}n_{i}$ is the number of units in stratum $j$ and, therefore,  $n=m_{s}=\prod_{i=1}^{s}n_{i}$ is the total number of runs.

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.85]{Hasse.jpg}      %width=\textwidth
\caption{Hasse diagram for factors}
\label{Fig::Hasse}
\end{center}
\end{figure}

Recall the expression for the linear polynomial hierarchical model from (\ref{eq::intro_ms}):
\begin{equation}
\label{eq::back_ms}
\bm{Y}=\bm{X}\bm{\beta}+\sum_{i=1}^{s}\bm{Z}_{i}\bm{\varepsilon}_{i},
\end{equation} 
where the first part $\bm{X\beta}$ contains the fixed effects, and $\sum_{i=1}^{s}\bm{Z}_{i}\bm{\varepsilon}_{i}$ stands for the random part, i.e.~it comprises information on variation occurring at each level of randomisation. All of the errors are assumed to be independent, having zero means and constant variances $\sigma^2_{i}$ and, as seen from the model formulation, additive.

In this case the fixed effects coefficients, $\bm{\beta}$ are usually estimated using the Generalised Least Square (GLS) formula:
\begin{equation}
\label{eq::back_gls}
\bm{\hat{\beta}}=(\bm{X}'\bm{VX})^{-1}\bm{X}'\bm{V}^{-1}Y,
\end{equation}
so that the variance of the estimators is
\begin{equation}
\label{eq::back_glsvar}
Var(\bm{\hat{\beta}})=(\bm{X}'\bm{VX})^{-1},
\end{equation}
where $\bm{V}$ is the variance-covariance matrix of the (normally distributed) responses, has a block-diagonal structure, and can be presented as:
\begin{equation}
\label{eq::back_glsV}
\bm{V}=\sum_{i=1}^{s}\sigma^2_{i}\bm{Z}_{i}\bm{Z}'_{i}=\sigma^{2}_{s}\left(\bm{I}_{n}+\sum_{i=1}^{s-1}\eta_{i}\bm{Z}_{i}\bm{Z}'_{i}\right).
\end{equation}
Here $\eta_{i}=\sigma^{2}_{i}/\sigma^{2}_{s}$ is defined as a variance ratio, denoting the magnitude of variability occurring at higher strata scaled with respect to the between-run variance.

At the analysis stage, as the variance components are unknown, their estimates are obtained and substituted in (\ref{eq::back_glsvar}) to obtain an estimated variance-covariance matrix of the parameters' estimators, $\bm{\hat{V}(\hat{\beta})}$; the estimation methods are discussed in more detail in Chapter \ref{ch::mse_ms}.

However, when an experiment is being planned, and none of the $\eta_{i}$ or $\sigma^2_{s}$ are known, there are two main approaches to deal with it. The first one is to search for designs assuming some point prior values of the $\eta$ parameter and, for these, evaluate the variance-based criteria, considering all strata simultaneously.

The case of split-plot experiments, with two strata, $\sigma^2_{s}=\sigma^2$ and $\eta=\sigma^2_{1}/\sigma^2$, has been extensively considered in the literature. \cite{Goos2001Doptimal} considered the three cases when D-optimal designs for split-plot experiments do not depend on the value of $\eta$; in other practical cases an estimate of the variance ratio is to be provided. \cite{Goos2003Doptimal} and \cite{Jones2007candidate} developed algorithms for finding $D$-optimal split-plot designs and for some examples demonstrated the robustness to the different values of $\eta$. An algorithm for constructing $D$-optimal split-split-plot designs can be found in the paper by \cite{Jones2009Doptimal}. 
%\cite{Goos2007tailor} -- split-plot for mixture experiments

The Bayesian alternative to the `exhaustive search' approach across the unknown parameter space allows specifying a prior distribution on the unknown parameters, and then evaluating the criteria by integrating over that prior. \cite{Arnouts2012staggered} presented a coordinate-exchange algorithm for constructing D-optimal designs for a staggered experimental structure (i.e. values of hard-to-change factors' are changed at different time points), with log-normal prior distributions being put on the variance ratios. Later \cite{Arnouts2015staggered} studied the staggered structured designs in the context of response surface modelling, and considered a few examples of D- and I-optimal designs.

\cite{Mylona2014optimal} introduced a composite criterion, combining $D$-optimality for the fixed and variance components; in such an approach both the criterion formula and the form of the prior distributions define the way of evaluation the criterion function, which often leads to the necessity of choosing a computationally efficient numerical methodology. \cite{Gilmour2009analysis} proposed an appropriate Bayesian analysis strategy of data from multistratum experiments, that is robust to designs non-orthogonality and is shown to be more reliable than REML approach.

However, in the general case a lot of computational effort might be required in order to be confident in the goodness of the design obtained by choosing from a set of assumed possible values of $\eta_i$, either using point priors or adapting general Bayesian strategy, especially when there are more than two strata and, therefore, the range of unknown parameters becomes multi-dimensional. The stratum-by-stratum approach (as first suggested by \cite{Trinca2001multistratum} and more recently improved by \cite{Trinca2015improved}) implies going from the highest stratum to the lowest, at each level choosing the set of treatments, and the units of higher level being treated as fixed block effects. Such methodology allows `protecting' against the case of large higher level variances, and eliminates the need for any prior information on unknown parameters. The authors later adapted the stratum-by-stratum  strategy for the inference criteria \citep{Trinca2016SPinference}, providing tools for constructing efficient designs for relatively small experiments in the presence of restricted randomisation.
\section{Acknowledgements}

>>>>

\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{tikz}
\usepackage{float}
%\usepackage{slashbox}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsfonts}
\usepackage{enumerate}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{bm}
\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}

@@

\begin{document}
\tableofcontents

\newpage
\section{Abstract}

@@

\begin{Large}

\end{Large}

@@

\subsection{Error estimation}
The error variance estimation can be obtained in two ways. The first one is the mean square error: $\hat{\sigma}^2_{mse}=\mbox{Residual SS}/(n-p)$, i.e. the residual sum of squares divided by the corresponding number of degrees of freedom (obtained from the ANOVA decomposition, the details can be found in \cite{Draper1998}). This estimator obviously depends on the model and on the number of its parameters. The other one is model-independent `pure' error, derived from the further decomposition of the residual sum of squares into the `pure' error and `lack-of-fit' components: $\hat{\sigma}^2_{PE}=\mbox{Pure error SS}/(n-t)$, where $t$ is the number of unique treatments applied and $d=n-t$ is the pure error degrees of freedom, i.e. the number of replicated points. In other words, the error is estimated by fitting the full treatment model:

@@

?? Adding background on MS experiments: briefly about the structure, pure-error reml, note on degresof freedom.

@@

\section{Individual criteria construction}
\label{ch::compound_criteria}
\input{compound_criterion}

\section{Multistratum Experiments}
\label{seq::ms_experiments}
\input{ms_experiments}

@@

%% Criteria for blocked experiments as a special case of MS
\subsection{Criteria for blocked experiments}
\label{subsec::compound_blocked}
\input{compound_blocked}

\subsection{Example. Case study}
\label{subsec::case_study}
\input{blocked_case_study}

\section{Future work}

\section{Acknowledgements}

### 2021-04-19 compound_criterion.tex
Initial commit

Standard design optimality theory is developed under the assumption that the fitted model represents the true relationship of interest and that there is no misspecification. However, it is obviously quite a strong belief and in reality we need to take into account at least the possibility that the model might not provide a good fit to the data. In Section \ref{sec::back_misspecification} a brief overview of various types of model misspecification is presented.

In this chapter we consider the case when the fitted polynomial model is nested within a larger model that is assumed to provide a better fit for the data:
\begin{equation}
\label{eq::full_model}
\bm{Y}=\bm{X}_1\bm{\beta}_1+\bm{X}_2\bm{\beta}_2+\bm{\varepsilon},
\end{equation}
where $\bm{X}_2$ is an $n\times q$ is an extension of the initial model matrix representing those extra $q$ potential terms that might represent the fitted model disturbance. The vector $\bm{\beta}_2$ denotes the corresponding parameters. This larger model has $p+q$ parameters in total and not all of them are estimable when the experiment is relatively small: $n<p+q$; this is the case we consider here.
\subsection{Lack-of-fit criterion}
We consider the Generalised criterion presented by Goos et al.: [more details?]
\begin{equation*}
\mbox{minimise }|\bm{X}'_{1}\bm{X}_1|^{-\frac{\kappa_{D}}{p}}\times \left|\bm{L}+\frac{\bm{I}_{q}}{\tau^{2}}\right|^{-\frac{\kappa_{LoF}}{q}} \times |\bm{A}'\bm{A}+\bm{I}_{q}|.^{\frac{\kappa_{bias}}{q}},
\end{equation*}
where $\bm{L}$ is known in literature as dispersion matrix, and the matrix $\bm{L}+\bm{I}_{q}/\tau^{2}$ in the second component is essentially the inverse posterior variance-covariance matrix of the vector of potential terms (up to a factor of $\sigma^2$).

@@

A diffuse prior shall be put on primary terms (as was done by \cite{DuMouchel1994}) -- an arbitrary mean and a variance going to infinity, and the prior on potential terms is the one specified before: $\bm{\beta}_2\sim\mathcal{N}(0,\bm{\Sigma}_{0})$,
$\bm{\Sigma}_{0}=\sigma^{2}\tau^{2}\bm{I}_{q}$, then the posterior
distribution of the coefficients is (as in \cite{Koch2007introduction} and \cite{DuMouchel1994}):
$$\bm{\beta}|\bm{Y}\sim \mathcal{N}(\bm{b},\bm{\Sigma}),\mbox{ where }\bm{b}=\bm{\Sigma X}'\bm{Y}, \bm{X}=[\bm{X}_1, \bm{X}_2],$$
$$\bm{\Sigma}=\left[\frac{\bm{K}}{\sigma^{2}\tau^{2}}+\sigma^{-2}(\bm{X'X})\right]^{-1}=\sigma^{2}[\bm{K}/\tau^{2}+\bm{X'}\bm{X}],^{-1}$$
and 
\begin{equation*}
\bm{K}=\begin{pmatrix}
\bm{0}_{p\times p} & \bm{0}_{p\times q}\\
\bm{0}_{q\times p} & \bm{I}_{q\times q}
\end{pmatrix}.
\end{equation*}
The marginal posterior of $\bm{\beta}_2$ is $\mathcal{N}(\bm{b}_{2},\bm{\Sigma}_{22})$, so using the general formula for the inverse of a block matrix
$$\begin{bmatrix}
 \bm{A}& \bm{B}\\
 \bm{C}& \bm{D}
\end{bmatrix}^{-1}=\begin{bmatrix}
\ldots & \ldots\\
\ldots & (\bm{D}-\bm{CA}^{-1}\bm{B})^{-1}
\end{bmatrix},$$
we can obtain the expression for $\bm{\Sigma}_{22}$:
\begin{align*}
[\bm{K}/\tau^{2}+\bm{X}'\bm{X}]^{-1}&=\begin{bmatrix}
 \bm{X}'_1\bm{X}_1& \bm{X}'_1\bm{X}_2 \\
 \bm{X}'_2\bm{X}_1& \bm{X}'_2\bm{X}_2+\bm{I}_{q}/\tau^{2}
\end{bmatrix}^{-1}\\&=\begin{bmatrix}
\ldots & \ldots\\
\ldots &
(\bm{X}'_2\bm{X}_2+\bm{I}_{q}/\tau^{2}-\bm{X}'_2\bm{X}_1(\bm{X}'_1\bm{X}_1)^{-1}\bm{X}'_1\bm{X}_2)^{-1}
\end{bmatrix},\\
\bm{\Sigma}_{22}&=\sigma^{2}[(\bm{K}/\tau^{2}+\bm{X}^{'}\bm{X})^{-1}]_{22}=\sigma^{2}\left(\bm{L}+\frac{\bm{I}_{q}}{\tau^{2}}\right)^{-1},
\end{align*} 
and $$(\bm{\Sigma}_{22})^{-1}=\frac{1}{\sigma^{2}}\left(\bm{L}+\frac{\bm{I}_{q}}{\tau^{2}}\right).$$

@@

$$(\bm{\beta}_{2}-\bm{b}_{2})^{'}(\bm{L}+\bm{I}_{q}/\tau^{2})(\bm{\beta}_{2}-\bm{b}_{2})\leq qs^{2}F_{q,\nu;1-\alpha},$$
where $\mathrm{F}(q,\nu,\alpha)$ is the $\alpha$-quantile of F-distribution with $q$ and $\nu$ degrees of freedom.\\

@@

\subsubsection{Point-based estimate}

One of the alternatives to the MC approach we would be willing to consider is to use the point prior for $\bm{\beta}_2$, that is $\bm{\beta}_2=\sigma\tau\bm{1}_q$ (where $\bm{1}_q$ is a $q$-dimensional vector with all components being equal to $1$), which is the standard deviation  of the initial (normal) prior. Therefore, $\bm{\tilde{\beta}}_2=\tau\bm{1}_q$, and the MSE(D)-part is
\begin{align*}
&|\bm{X}'_{1}\bm{X}_{1}|^{-1}(1+\bm{\tilde{\beta}}_{2}'\bm{X}_2^{'}\bm{X}_1\bm{M}^{-1}\bm{X}_1^{'}\bm{X}_2\bm{\tilde{\beta}}_{2})\\=&|\bm{X}'_{1}\bm{X}_{1}|^{-1}(1+\tau^2\bm{1}'_q\bm{X}_2^{'}\bm{X}_1\bm{M}^{-1}\bm{X}_1^{'}\bm{X}_2\bm{1}_q)\\=&|\bm{X}'_{1}\bm{X}_{1}|^{-1}\left(1+\tau^2\sum_{i,j=1}^{q}[\bm{X}_2^{'}\bm{X}_1\bm{M}^{-1}\bm{X}_1^{'}\bm{X}_2]_{(i,j)}\right),
\end{align*}
so essentially the last part of the equation is the summation of matrix elements, which obviously takes less computational time even for large dimensions $q$.

>>>>

Standard design optimality theory is developed under the assumption that the fitted model represents the true relationship of interest and that there is no misspecification. However, it is obviously quite a strong belief and in reality we need to take into account at least the possibility that the model might not provide a good fit to the data. 
%In Section \ref{sec::back_misspecification} a brief overview of various types of model misspecification is presented. 
?? Other types of model misspecification (in background)

In this chapter we consider the case when the fitted polynomial model is nested within a larger model that is assumed to provide a better fit for the data:
\begin{equation}
\label{eq::full_model}
\bm{Y}=\bm{X}_1\bm{\beta}_1+\bm{X}_2\bm{\beta}_2+\bm{\varepsilon},
\end{equation}
where $\bm{X}_2$ is an $n\times q$ is an extension of the initial model matrix representing those extra $q$ potential terms that might represent the fitted model disturbance. The vector $\bm{\beta}_2$ denotes the corresponding parameters. This larger model has $p+q$ parameters in total and not all of them are estimable when the experiment is relatively small: $n<p+q$; this is the case we consider here.
\subsection{Lack-of-fit criterion}
%We consider the Generalised criterion presented by Goos et al.: [more details?]
%\begin{equation*}
%\mbox{minimise }|\bm{X}'_{1}\bm{X}_1|^{-\frac{\kappa_{D}}{p}}\times \left|\bm{L}+\frac{\bm{I}_{q}}{\tau^{2}}\right|^{-\frac{\kappa_{LoF}}{q}} \times |\bm{A}'\bm{A}+\bm{I}_{q}|.^{\frac{\kappa_{bias}}{q}},
%\end{equation*}
%where $\bm{L}$ is known in literature as dispersion matrix, and the matrix $\bm{L}+\bm{I}_{q}/\tau^{2}$ in the second component is essentially the inverse posterior variance-covariance matrix of the vector of potential terms (up to a factor of $\sigma^2$).

@@

To be able to control the scope of the potential terms, we adapt Bayesian approach:  diffuse prior shall be put on primary terms (as was done by \cite{DuMouchel1994}) -- an arbitrary mean and a variance going to infinity, and the prior on potential terms is a normal distribution: $\bm{\beta}_2\sim\mathcal{N}(0,\bm{\Sigma}_{0})$, with the prior variance scaled with respect to the error variation:
$\bm{\Sigma}_{0}=\sigma^{2}\tau^{2}\bm{I}_{q}$. Then the posterior
distribution of the coefficients is (as in \cite{Koch2007introduction} and \cite{DuMouchel1994}):
$$\bm{\beta}|\bm{Y}\sim \mathcal{N}(\bm{b},\bm{\Sigma}),\mbox{ where }\bm{b}=\bm{\Sigma X}'\bm{Y}, \bm{X}=[\bm{X}_1, \bm{X}_2],$$
$$\bm{\Sigma}=\left[\frac{\bm{K}}{\sigma^{2}\tau^{2}}+\sigma^{-2}(\bm{X'X})\right]^{-1}=\sigma^{2}[\bm{K}/\tau^{2}+\bm{X'}\bm{X}],^{-1}$$
and 
\begin{equation*}
\bm{K}=\begin{pmatrix}
\bm{0}_{p\times p} & \bm{0}_{p\times q}\\
\bm{0}_{q\times p} & \bm{I}_{q\times q}
\end{pmatrix}.
\end{equation*}
The marginal posterior of $\bm{\beta}_2$ is $\mathcal{N}(\bm{b}_{2},\bm{\Sigma}_{22})$, so using the general formula for the inverse of a block matrix
$$\begin{bmatrix}
 \bm{A}& \bm{B}\\
 \bm{C}& \bm{D}
\end{bmatrix}^{-1}=\begin{bmatrix}
\ldots & \ldots\\
\ldots & (\bm{D}-\bm{CA}^{-1}\bm{B})^{-1}
\end{bmatrix},$$
we can obtain the expression for $\bm{\Sigma}_{22}$:
\begin{align*}
[\bm{K}/\tau^{2}+\bm{X}'\bm{X}]^{-1}&=\begin{bmatrix}
 \bm{X}'_1\bm{X}_1& \bm{X}'_1\bm{X}_2 \\
 \bm{X}'_2\bm{X}_1& \bm{X}'_2\bm{X}_2+\bm{I}_{q}/\tau^{2}
\end{bmatrix}^{-1}\\&=\begin{bmatrix}
\ldots & \ldots\\
\ldots &
(\bm{X}'_2\bm{X}_2+\bm{I}_{q}/\tau^{2}-\bm{X}'_2\bm{X}_1(\bm{X}'_1\bm{X}_1)^{-1}\bm{X}'_1\bm{X}_2)^{-1}
\end{bmatrix},\\
\bm{\Sigma}_{22}&=\sigma^{2}[(\bm{K}/\tau^{2}+\bm{X}^{'}\bm{X})^{-1}]_{22}=\sigma^{2}\left(\bm{L}+\frac{\bm{I}_{q}}{\tau^{2}}\right)^{-1},
\end{align*} 
and $$(\bm{\Sigma}_{22})^{-1}=\frac{1}{\sigma^{2}}\left(\bm{L}+\frac{\bm{I}_{q}}{\tau^{2}}\right).$$

@@

$$(\bm{\beta}_{2}-\bm{b}_{2})^{'}(\bm{L}+\bm{I}_{q}/\tau^{2})(\bm{\beta}_{2}-\bm{b}_{2})\leq qs^{2}F_{q,\nu;1-\alpha},$$
where $\mathrm{F}(q,\nu,\alpha)$ is the $\alpha$-quantile of F-distribution with $q$ and $\nu$ degrees of freedom. Then, similarly, to $DP$-optimality, minimsing the volumeof this confidence region is equivalent to minimising $\left|\bm{L}+\frac{\bm{I}_{q}}{\tau^{2}}\right|^{-1/q}F_{q,d;1-\alpha_{LoF}}$. The matrix is directly related to the lack-of-fit component in the Generalised criteria developed by Goos et al. (2005) -- in case the error estimate does not depend on the design.

@@

\subsubsection{Point-based estimate}
?? Should this Section be much shorter
One of the alternatives to the MC approach we would be willing to consider is to use the point prior for $\bm{\beta}_2$, that is $\bm{\beta}_2=\sigma\tau\bm{1}_q$ (where $\bm{1}_q$ is a $q$-dimensional vector with all components being equal to $1$), which is the standard deviation  of the initial (normal) prior. Therefore, $\bm{\tilde{\beta}}_2=\tau\bm{1}_q$, and the MSE(D)-part is
\begin{align*}
&|\bm{X}'_{1}\bm{X}_{1}|^{-1}(1+\bm{\tilde{\beta}}_{2}'\bm{X}_2^{'}\bm{X}_1\bm{M}^{-1}\bm{X}_1^{'}\bm{X}_2\bm{\tilde{\beta}}_{2})\\=&|\bm{X}'_{1}\bm{X}_{1}|^{-1}(1+\tau^2\bm{1}'_q\bm{X}_2^{'}\bm{X}_1\bm{M}^{-1}\bm{X}_1^{'}\bm{X}_2\bm{1}_q)\\=&|\bm{X}'_{1}\bm{X}_{1}|^{-1}\left(1+\tau^2\sum_{i,j=1}^{q}[\bm{X}_2^{'}\bm{X}_1\bm{M}^{-1}\bm{X}_1^{'}\bm{X}_2]_{(i,j)}\right),
\end{align*}
so essentially the last part of the equation is the summation of matrix elements, which obviously takes less computational time even for large dimensions $q$.

### 2021-04-19 thesis_main/Chapter5_Examples.tex
Initial commit

Here we will study the optimal designs in terms of the criteria (\ref{eq::MSE_D}), (\ref{eq::MSE_D_point}) and (\ref{eq::MSE_L}) (referred to as $MSE(D)$, $MSE(D)P$ and $MSE(L)$ criteria respectively) in the framework a factorial experiment, with $5$ factors, each is at three levels. The small number of runs ($40$) allows estimation of the full-second order polynomial model ($p=21$), but we assume that the `true' model contains also all third-order terms (linear-by-linear-by-linear and quadratic-by-linear interactions), $q=30$ of them in total.. 
[Some comparison with other criteria is to be added]

The intercept is a nuisance parameter, and so the criteria are adapted in such a way that the $DP$- and $LP$-components are replaced by $DPs$ and $LPs$. After considering the mean square error matrix (\ref{eq::MSE}) for $p-1$ parameters, in the MSE-based components the full information matrix is replaced by the one excluding the intercept -- $\bm{X}'_{1}\bm{Q}_0\bm{X}_{1}$ (see (\ref{eq::s_infmatrix})), where $\bm{X}_1$ is the model matrix without the intercept, $\bm{Q}_0=\bm{I}_n-\frac{1}{n}\bm{11}'$, and $\bm{\tilde{M}}$ is the full information matrix for the model with the intercept. Otherwise, the procedure of obtaining the $MSE(D)$- and $MSE(L)$-based components remains the same.

>>>>

%%% Mean Square Error based criteria, Example
Here we will study the optimal designs in terms of the criteria (\ref{eq::MSE_D}), (\ref{eq::MSE_D_point}) and (\ref{eq::MSE_L}) (referred to as $MSE(D)$, $MSE(D)P$ and $MSE(L)$ criteria respectively) in the framework of the experiment described before: five three-level factors, the full second-order fitted polynomial model ($p=21$), potential terms comprising all quadratic-by-linear and linear-by-linear-by-linear terms, $q=30$ of them in total.

As before, the intercept is a nuisance parameter, and so the criteria are adapted in such a way that the $DP$- and $LP$-components are replaced by $DPs$ and $LPs$, as in (\ref{eq::DPs_inter}) and (\ref{eq::LPs_inter}). After considering the mean square error matrix (\ref{eq::MSE}) for $p-1$ parameters, in the MSE-based components the full information matrix is replaced by the one excluding the intercept -- $\bm{X}'_{1}\bm{Q}_0\bm{X}_{1}$ (see (\ref{eq::s_infmatrix})), where $\bm{X}_1$ is the model matrix without the intercept, $\bm{Q}_0=\bm{I}_n-\frac{1}{n}\bm{11}'$, and $\bm{\tilde{M}}$ is the full information matrix for the model with the intercept. Otherwise, the procedure of obtaining the $MSE(D)$- and $MSE(L)$-based components remains the same.

### 2021-04-19 thesis_main/Chapter6_Blocked.tex
Initial commit

@@

@@

>>>>

In the previous chapter we introduced the compound criteria comprising minimisation of the model estimates' variation and bias arising from the potential model misspecification of a known form, together with the minimising the associated lack-of-fit. We considered an example of a factorial experiment and examined how the criteria would work in terms of the resulting optimal designs and the relationships between the individual components in terms of the relationships between the weight allocations and designs' performances.

A natural extension to be implemented in this chapter is to adapt the developed criteria to an other widely used experimental  framework, blocked experiments. Here we also consider a real-life example of a blocked experiment and provide some practical solutions for the experimenters according to the particular case-specific aims, objectives and restrictions.

\section{MSE-Based Criteria Adapted for Blocked Experiments}

@@

\subsection{Trace-based criterion}

@@

\subsection{Example}
\input{Chapter6_Example}

\newpage
\section{Case Study Experiment}
\label{sec::case_study}
\input{Chapter6_CaseStudy}

%% Discussion points
\section{Conclusions}

In general, this chapter is about exploring the flexibility of the design optimality criteria and implementation introduced previously. We adapted the approach to the framework of a blocked experiment, and by considering an example confirmed some of the key features of the relationships between the individual components, and tendencies in optimal designs obtained with various weight allocation schemes. Due to the relatively high time consumption of the determinant-case criterion calculation, the simplest suggested alternative seemed quite promising in terms of providing adequate designs in a much more reasonable time.

The second part of the chapter is devoted to a real-life design problem, such that the given framework of a blocked factorial experiment with some suspicion regarding possible model disturbance suggested applying the MSE-based criterion for the design construction. In addition, the structure of the criterion and the search algorithm implementation allowed for necessary adaptation with regard to the pre-specified restrictions, so that a set of possible solutions was presented satisfying the objectives of the particular experiment.

### 2021-04-19 thesis_main/Chapter6_CaseStudy.tex
Initial commit

A company specialising in food supplements production for some animals wanted to conduct an experiment with the main aim of figuring out whether a slight decrease in the recommended dosages of three particular products taken together would make a significant impact on the resulting ``performance'', which is expressed in terms of some continuous response. The dosage range of interest for each supplement is from $90\%$ to $100\%$ of the standard recommendation; it is desired that there would be $3$ levels (i.e. taking the values of $90\%$, $95\%$ and $100\%$). Carrying out the experiment with more than $3$ levels was possible but more complicated, as measuring, for example, $92.5\%$ of the recommended dosage might be quite tricky. However, the search was performed among a larger $5$-level candidate set of points (which, obviously, contains all $3$-level design points), but due to the form of the fitted model and the criteria used, the resulting designs' points consist of only $3$-level points.

>>>>

%% Case-study, experimantal layout and results

In this section we will consider a real-life example of a design problem where the criterion derived in this chapter was used to produce an optimal design corresponding to the objectives and the framework of the experiment. We will present several alternative solutions, so that the experimenters will be able to make the final choice of the design.

Due to the confidentiality requirements all particular details of the experiments cannot be disclosed, and the description of the experimental layout is to be anonymised.

\subsection{The experimental framework, aims and limitations}
A company specialising in food supplements production for some animals wanted to conduct an experiment with the main aim of figuring out whether a slight decrease in the recommended dosages of three particular products taken together would make a significant impact on the resulting ``performance'', which is expressed in terms of some continuous response. The dosage range of interest for each supplement is from $90\%$ to $100\%$ of the standard recommendation; it is desired that there would be $3$ levels (i.e. taking the values of $90\%$, $95\%$ and $100\%$). Carrying out the experiment with more than $3$ levels was possible but more complicated, as measuring, for example, $92.5\%$ of the recommended dosage might be quite tricky. However, the search was performed among a larger $5$-level candidate set of points (which, obviously, contains all $3$-level design points), but due to the form of the fitted model and the criteria used, the resulting designs' points consist of only $3$-level points.
